---
title: "Leakage-aware biomarker panel discovery"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Leakage-aware biomarker panel discovery}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

## Why nested validation is necessary

A biomarker workflow often ranks all markers, chooses a candidate pool, searches
many panels, and then reports cross-validation for the selected panel. That
procedure leaks outcome information because the assessment observations have
already influenced feature selection.

`detectPanel` places the following operations inside each outer analysis set:

1. detectability and stability scoring;
2. individual marker AUC estimation;
3. candidate filtering and ranking;
4. panel combination search on shared inner splits;
5. scaling, logistic fitting (with deterministic ridge fallback for unstable separation-prone fits), and threshold selection.

The outer assessment set is used only for prediction.

## Synthetic example

```{r example, eval=FALSE}
library(detectPanel)

set.seed(9)
n <- 36
p <- 12
y <- rep(c("Control", "Case"), each = n / 2)
counts <- matrix(
  rpois(p * n, 50), nrow = p,
  dimnames = list(paste0("m", seq_len(p)), paste0("s", seq_len(n)))
)
counts[1:3, y == "Case"] <- counts[1:3, y == "Case"] + 60
meta <- data.frame(group = y, row.names = colnames(counts))

fit <- discover_panel(
  counts, meta,
  outcome = "group",
  positive = "Case",
  candidate_n = 7,
  min_mean = 5,
  min_median = 2,
  min_detection = 0.3,
  min_group_detection = 0.2,
  min_auc = 0.55,
  outer_v = 3,
  outer_repeats = 1,
  inner_v = 3,
  inner_repeats = 1
)

fit$nested$outer_summary
fit$nested$feature_frequency
```

## Interpreting the two models

The nested predictions estimate internal generalization performance. After that
assessment, `discover_panel()` refits the most frequently selected exact panel
on all samples. This final model is useful for locked prediction on a new cohort,
but its own training AUC is not an independent validation result.

## Candidate fallback

By default, a training split fails when fewer than `panel_size` markers satisfy
the prespecified thresholds. This makes threshold violations visible. During
exploratory work, `allow_fallback = TRUE` may be used to fill the pool with the
highest-ranked finite markers. The `used_fallback` field must then be reviewed
and reported.
